To Airbnb?: A look at Long vs Shorm Term Leasing

### 2018-11-04 papers/full_paper.Rmd
Update to geographic sample, models, host, etc.  Rough draft.

---
title: "To Airbnb?: A look at Long vs Shorm Term Leasing"
subtitle: "(Andy Krause and Gideon Aschwanden)Rev"
author: "Greenfield Advisors and University of Melbourne"
date: "Revised: October 16th, 2018"
output:
  pdf_document: default
  html_document: default
header-includes: 
   - \usepackage{setspace}\doublespacing
   - \usepackage{caption}
   - \captionsetup[table]{skip=10pt}
geometry: margin=3.3cm
bibliography: abb.bib
editor_options: 
  chunk_output_type: console
---

@@

  library(tidyverse)
  library(sf)
  library(geosphere)
  library(ggmap)
  library(multidplyr)
  library(tmap)
  library(xtable)
  library(ROCR)
  library(xtable)

@@

```
## Abstract

@@

To address these questions and to provide some guidance to potential and current investors and owners, we examine the short- and long-term residential rental markets in Melbourne, Australia. Melbourne has a thriving short-term rental market with over 20,000 total properties listed on the Airbnb portal, the source of our short-term market data. To avoid confounding market dynamics caused by traditional vacation rentals (beach, ski and other recreation) we limit our study area to metropolitan area and suburban/rural areas lacking these recreational amenities. Additionally, for any individual property the choice between short and long term leasing may be influenced by individual financial situations (tax and mortgage financing) as well as non-financial concerns such as time availability and personal attachement to the property.  While we analyze individual properties, we acknowledge that there are many unobserved condiseration and, therefore, we generalize our findings to market segments, locations and host policy decisions.

@@

We begin by filtering observations by time. Our hypothetical leasing strategy decision is that of an investor who purchases a property on September 1, 2015 and must decide on traditional long-term rental versus an Airbnb-type, short-term rental approach to generating income. To determine the likely LTR nightly rate for any given property, we will use the observed rental transactions from September 1, 2014 to August 31, 2015. Taking the Airbnb or STR approach allows the owner to change rates over time and is subjected to daily changes in the market (supply and demand factors). In this simulated example we will use Airbnb data from September 1, 2015 to August 31, 2016 to represent the actual Airbnb market conditions over the time period in question.

@@

@@

### Rates and Rental Values

@@

### Location and Submarkets

@@

It is likely that the short-term rental market follows a similar market hierarchy, however, we can imagine that a few different spatial features can influence the short-term market.  While there are many possible additional features, we consider two to be the most likely to influence rates or occupancy: 1) Proximity to beaches; and 2) Proximity to key tourist activities and events. Using the three broad submarkets from the long-term market (Inner, Middle and Outer) we have created a five sub-market system as a starting point for spatially analyzing short and long-term rentals in Melbourne:

1. City-Core (Select inner suburbs with tourist activities and near CBD)
2. City (Inner Suburbs, not Beach, not Core)
3. Suburban (Middle suburbs, not Beach)
4. Rural (Outer suburbs, not Beach)
5. Beach (Properties within 500m of Port Phillip Bay east of Yarra River)

@@

We assign these designations by: 1) Adding suburb designations to the properties; 2) Assigning submarkets 1-4 based on suburb location; and 3) Indicating proximity to beach and labeling as 'Beach' submarket.  This process is repeated for both the short- and long-term data. Finally, we filter our data to observations in suburbs that contain at least three STR and LTR observations. The location of the short- and long-term rentals, colored by submarket, are shown in Figures 2 and 3.

@@

To better visualize the relative frequency of the submarkets in the two tenure types we have broken down the data by submarket by property type by tenure type. A number of interesting observations are illustrated here (Figure 4). First, the vast majority of Airbnb units are apartments in city-core or city locations. As expected and as shown by the map (Figure 2), the long-term properties are more evenly spread across the metro region than the short-term properties.  Long-term houses are more peripheral than apartments, a result of urban development patterns and land economics. Beach properties are, relatively speaking, over-represented in the short-term market, especially in apartments, while suburban and rural apartments, somewhat common in long-term markets, are nearly absent in the short-term market. Overall, the suburban and rural Airbnb markets are quite thin, later results focusing on these markets will need to be examined in light of this small sample size.

@@

### Owner Types

@@

First, we must deal with those properties that were not listed on Airbnb for the entire 12-month period.  Our data cleaning eliminated all properties that were listed for less than six months, however, it would be incorrect to compare the revenue of a property listed on Airbnb for seven months with the hypothetical annual revenue from a long-term rental.

Second, many of the short-term properties have a significant number of blocked days, days in which the property was not offered on Airbnb.  Here too, there is a problem with comparing observed revenues from a property with a 40\% block rate with a hypothetical year long revenue stream, especially if the property was leased out on other portals during those blocked days.  And, even if we knew that the blocked days were simply due to the host wanting to use the property themselves during that time, this deduction in revenue should not necessarily apply to the hypothetical investor in our examined scenario.  In short, in either case, the actual observed short-term revenue would be biased on the low end compared to what a revenue maximizing owner would be likely able to accomplish.

As a result, we have calculated three separate revenues values for the short-term properties:

1. Actual ($REV_a$)}: Observed bookings multiplied by median nightly rate
2. Likely ($REV_l$)}: Observed bookings extrapolated to a full year representing the likely actual revenue if the property had been listed all year long. The additional bookings are imputed by comparing the property specific performance during the listed days to overall market performance and then applying that relative ratio to the overall market performance on the non-list days.  
3. Potential ($REV_p$)}: Observed bookings plus imputed bookings during blocked days.  The imputed bookings are imputed by comparing the property specific performance during the non-blocked days to overall market performance and then applying that relative ratio to the overall market performance on the blocked days.

As the 'Actual' calculation systematically under counts the revenue from properties that were not on the Airbnb portal for the entire year we disregard it going forward and consider the 'Likely' revenue metric to be a better estimate of market performance. We chart the distribution of the Likely ($REV_l$) and Potential ($REV_p$) short-term revenues against a distribution of the observed long-term revenues for the long-term properties (Figure 5). Revenues from the long-term rentals are far more consistent, with the vast majority totaling between \$12,000 to \$30,000 per year.  Likely short-term revenues range from neglible to highs of greater than \$80,000, with many of the properties showing very limited (less than \$10,000) revenue. The range of the distribution for the Potential short-term revenue is similar, however, the distribution itself is much flatter, with many more properties showing potential incomes in the \$20,000 to \$40,000 range.

While this simple analysis does not control for location and product differences, it does highlight the fact that some Airbnb owners are likely making greater revenues than in long-term rentals, but that most are likely not. The high number of very low yield Airbnb properties in the Likely condition suggests that some owners may not be focused on maximizing revenue, but instead prefer the flexibility of short-term leasing.^[Do note that we have not considered any overall market demand issues when analyzing potential revenue.]

@@

In this section we compare net revenues between long- and short-term rental options.  As mentioned above, there are no direct matches between the datasets. In other words, we have not directly observed a property which has been in a long-term rental followed by short-term rental or vice versa.  We will, instead, make the comparisons between the two by imputing likely long-term rental revenues for the observed short-term properties and then comparing the measured net revenues of the short-term properties to their estimated imputed long-term revenues. We first focus on using the Likely measure of net revenue for the short-term properties as discussed above.  We save discussion and analysis of the Potential revenue for the sensitivity tests later in the study.

We make our imputations via hedonic price models. We model the long-term market with the long-term data and then use this model to predict or impute the likely long-term weekly rental prices for the short-term properties. We use the following model specification:

@@

@@

Overall, 28.4% short-term rental properties were more profitable as a short-term rental than it would have been if rented long-term over the course of the year. While illustrative, a simple univariate analysis does not fully control for property characteristics or owner actions -- which are a key component of short-term rental success.  To better understand which factors are related to a higher likelihood of short-term preference, we develop a set of logistic regression models below.

@@

The results of all three models are shown in Table 4. For $PType$ Apartments are the reference category,  for $BedBath$, 1 Bed / 1Bath is the reference category, for $Submrkt$ the City-Core is the reference category and for $CancelPolicy$ Strict is the reference category.  As the additional variables are added, the model fit diagnostics steadily increase, suggesting that the increased complexity of the models adds to the quality of the fit.

In Model 1, we see that when controlling for structural characteristics, Houses have a significantly lower probability of being more profitable in short-term leasing than apartments. Additionally, 1 bed / 1 bath units show greater likelihood of short-term preference than 2 bed / 2 bath combinations. This could be due to the fact that 2 bed units may be very attractive in the long-term market as they are a useful product for two, non-family roomates to share. Larger units, 3 bedrooms plus, generally show no statistical difference in preference porbability to the 1 bed / 1 bath reference unit, with the exception of the 4 bed / 2 bath properties.

Adding the submarket variables (Model 2) shows that all areas have a much lower probability of short-term prefernce when compared to the City Core (the reference category) when controlling for structural characteristics.  The difference between the Rural and City-Core is not significant, however. The coefficients for the structural characteristics show the same signs and relatively small changes in magnitude from Model 1 to Model 2, speaking to the robustness of those estimates. The coefficient for House properties does decrease significantly in magnitude when adding the geographic submarkets.  This is likely due to the fact that there are sturctural differences between the geographic submarkets.  Houses, for example, are concentrated more heavily in City and Suburban areas than in the City-Core and some of the impact of House is being mediated by the geographic submarkets.

Finally, in Model 3 we add in host behavior or policies. Allowing a greater number of guests per bedroom results in a higher probability of short-term preference.  This variable may be operating as proxy for additional sleeping spaces which allow for a higher nightly rate, ceteris paribus, to be charged.  The greater the minimum stay, the lower the probability of short-term preference.  Longer minimum stays may equate to lower occupancy rates as one or two night travelers are exempted from the property.  Owners with longer minimum stays may be looking to minimize costs such as cleaning and administration.  Flexible cancellation policies result in a lower probability of short-term preference, while a Strict policy has a positive relationship with short-term preference. Strict cancellation may be a luxury for the best and highly desired properties on the market and, within our model, may be acting as a proxy for some unobserved measure of quality or demand while a flexible cancellation may be signalling the opposite.

@@

  m1 <- as.data.frame(summary(lik_glms$str)$coefficients)
  m2 <- as.data.frame(summary(lik_glms$loc)$coefficients)
  m3 <- as.data.frame(summary(lik_glms$host)$coefficients)

@@

 ##  Create the variable section of table
 full.table <- as.data.frame(cbind(rep('x', length(m1c)), m1c, m2c, m3c))
 names(full.table) <- c('Variable', 'Model 1', "Model 2", "Model 3")
 full.table$Variable <- c('Intercept', 'House (Type)', '2 Bed & 1 Bath',
                          '2 Bed & 2 Bath', '3 Bed & 1 Bath', '3 Bed & 2 Bath',
                          '3 Bed & 3 Bath','4 Bed & 2 Bath','4 Bed & 3 Bath', 
                          'City (Submarket)',
                          'Suburban (Submarket)', 'Rural (Submarket)',
                          'Beach (Submarket)', 'Guests per Bedroom',
                          'Minimum Stay', 'Flexible Cancel')

@@

  # Extract diagnostics
  str.dx <- logDx(lik_glms$str, str_tdf, str_tdf$lik_pref)
  sm.dx <- logDx(lik_glms$loc, str_tdf, str_tdf$lik_pref)
  host.dx <- logDx(lik_glms$host, str_tdf, str_tdf$lik_pref)

@@

  # Print
  print(x.full,
       include.rownames=FALSE,
       hline.after=c(-1,0,16,21,21),
       caption.placement='top')

@@

To test if the above results are influence by the assumptions we have made throughout this analysis, we perform two sets of sensitivity tests.  First, we examine how the results change if Potential revenue, not Likely revenue, of the short-term properties is used in the comparisons.  Second, we categorize the hosts, split the data and re-analyze the performance of properties based on hypothesized motivations and actions of the hosts.

#### Potential Revenue

The above uses the Likely^[Extrapolated out in cases of Airbnb properties on the market for less than a year] revenue in the comparisons with the imputed long-term revenue.  Doing so may underestimate the possible revenue -- what may matter most to the hypothetical buyer in our scenario -- due to not factoring in the impact of blocked days.  If we, instead, assume that blocked days could be leased out on the short-term market at the same overall occupancy rate (relative for each property and the market as a whole on those days) as the unblocked (observed) days we can produce a potential revenue figure. Note that it is possible (likely?) that the large oversupply of available units in the market that would occur if no properties blocked their units would decrease both prices and occupancy rates.  We do not attempt to correct for this likelihood, rather we offer these figures as high end estimates.

Using the potential revenue, 36.1% of all properties are short-term preferable.  Breaking that down by type and by geographic submarket (Table 3) shows similar findings, in terms of relative comparativeness, to the Likely revenue analysis above.  Apartments have a higher short-term preference than do Houses and City-Core properties dominate the other geographic regions.  One notable difference here is that the Suburban and Rural properties perform rather poorly under a potential revenue scenario, relatively speaking, suggesting that they may be operating near their peak in the current condition.

```{r pot_table, results='asis', echo=FALSE, message=FALSE, comment=FALSE}

 # Make Basic comparison table of type
  tpref.table <- abbCreateCompTable(ic.df=str_tdf,
                                    split.field='type')

 # Extract Potential Results
  tpref.table <- tpref.table[tpref.table$rev.type == 'Potential', ]

 # Make baseic comparison table of submarkets
  spref.table <- abbCreateCompTable(ic.df=str_tdf,
                                    split.field='geo_mrkt')

 # Extract potential results
  spref.table <- spref.table[spref.table$rev.type == 'Potential', ]

 # Combine and round
  st.table <- rbind(tpref.table, spref.table)
  st.table[,2] <- paste0(sprintf('%.1f', st.table[,2]*100), '%')

 ## Fix names and print

  names(st.table) <- c('Sub-Market', 'Short-Term Preferable')
  options(xtable.comment = FALSE)
  print(xtable(st.table[,1:2],
               caption='Potential Revenue Preferability\n',
               digits=1, align='llr'),
               include.rownames=FALSE,
               caption.placement='top',
               hline.after=c(-1, 0, 2, 7, 7))

Next, we re-run the logistic regression models using short-term preference in the potential case as the dependent variable. Generally speaking, the results from the new series of models (Table 4) are qualitatively similar to those derived in the Likely revenue case (Table 2).  The overall magnitude of the coefficient for House (vs Apartments) is lowered in the Potential case, suggesting that houses may be blocked off more that Apartments.  The product types (bed/bath) show similar trends, though with larger discounts for the 2 Bed / 1 Bath and 2 Bed/2 Bath units.  Likewise the geographic submarkets and host policies show similar effects in terms of direction of impact with some slight deviations in magnitude.

```{r pot_logmodel, echo=FALSE, message=FALSE, comment=FALSE}

  pm1 <- as.data.frame(summary(pot_glms$str)$coefficients)
  pm2 <- as.data.frame(summary(pot_glms$loc)$coefficients)
  pm3 <- as.data.frame(summary(pot_glms$host)$coefficients)

 ## Make Significance stars

  pm1sig <- makeStatSig(pm1[,4])
  pm2sig <- makeStatSig(pm2[,4])
  pm3sig <- makeStatSig(pm3[,4])

 ## Create Table coef vectors

 pm1$table.coef <- paste0(ifelse(pm1$Estimate>0, ' ', ''),
                         sprintf('%.3f', round(pm1$Estimate,3)),
                         pm1sig)
 pm2$table.coef <- paste0(ifelse(pm2$Estimate>0, ' ', ''),
                         sprintf('%.3f',round(pm2$Estimate,3)),
                         pm2sig)
 pm3$table.coef <- paste0(ifelse(pm3$Estimate>0, ' ', ''),
                         sprintf('%.3f',round(pm3$Estimate,3)),
                         pm3sig)

 ## Snap to same length

 pm1c <- c(pm1$table.coef, rep(' ', nrow(pm3)-nrow(pm1)))
 pm2c <- c(pm2$table.coef, rep(' ', nrow(pm3)-nrow(pm2)))
 pm3c <- pm3$table.coef

 ## Put together as variables table

 pfull.table <- as.data.frame(cbind(rep('x', length(pm1c)), pm1c, pm2c, pm3c))
 names(pfull.table) <- c('Variable', 'Model 1', "Model 2", "Model 3")
 pfull.table$Variable <- c('Intercept', 'House (Type)', '2 Bed & 1 Bath',
                          '2 Bed & 2 Bath', '3 Bed & 1 Bath', '3 Bed & 2 Bath',
                          '3 Bed & 3 Bath','4 Bed & 2 Bath','4 Bed & 3 Bath', 
                          'City (Submarket)',
                          'Suburban (Submarket)', 'Rural (Submarket)',
                          'Beach (Submarket)', 'Guests per Bedroom',
                          'Minimum Stay', 'Flexible Cancel')

  # Extract diagnostics
  pstr.dx <- logDx(pot_glms$str, str_tdf, str_tdf$pot_pref)
  psm.dx <- logDx(pot_glms$loc, str_tdf, str_tdf$pot_pref)
  phost.dx <- logDx(pot_glms$host, str_tdf, str_tdf$pot_pref)

  # Create diagnostic table
  pdiag.table <- data.frame(Variable=c('', 'Diagnostics', 'AIC', 'LogLik',
                                      'AUC'),
                           Str.Model=c('', '', round(pstr.dx$AIC,0),
                                       round(pstr.dx$logLik, 0),
                                       round(as.numeric(pstr.dx$auc@y.values),
                                             3)),
                           Sm.Model=c('', '', round(psm.dx$AIC,0),
                                      round(psm.dx$logLik, 0),
                                      round(as.numeric(psm.dx$auc@y.values),
                                            3)),
                           Host.Model=c('', '', round(phost.dx$AIC,0),
                                        round(phost.dx$logLik, 0),
                                        round(as.numeric(phost.dx$auc@y.values),
                                              3)))
 ## Fix names

  names(pdiag.table) <- c('Variable', 'Model 1', "Model 2", "Model 3")

 # Bind together

 pfull.table <- rbind(pfull.table, pdiag.table)

```{r table_potlogmodel, results='asis', echo=FALSE, message=FALSE, comment=FALSE}

 # XTable

 px.full <- xtable(pfull.table,
                   caption='Logistic Modeling Results (Potential Revenue)\n')
 align(px.full) <- 'lllll'

 # Print
 print(px.full,
       include.rownames=FALSE,
       hline.after=c(-1, 0, 16, 21, 21),
       caption.placement='top')

Overall, the choice of revenue type, Likely or Potential, does affect some of the conclusions to be derived from the results above.  As expected, under the Potential case there are more properties (36\% vs 28\%) that exhibit short-term preference.  Under the scenario, the possibility of increasing revenue by moving from long- to short-term leasing appears much better.  Again, however, we have not attempted to measure or estimate how this hypothetical increase of properties in the market (due to the elimation of blocked days) would impact overall market dynamics.  Suffice it to say that it would like drive down occupancies and maybe price, therefore leading to a decrease in short-term preference likelihood.

In explaining the drivers of short-term preference, the logistic regression models do not change materially with the switch in dependent variable.  This suggests that the underlying drivers of short-term preference remain relatively constant regardless of the choice of revenue estimate. Only the base probability of short-term preference increases in the potential revenue case.

More broadly speaking, we explain the differences between the two revenue estimates as follows:  The actual revenue case shows what is happening in the market while the potential revenue case highlights what a given property owner may be able to achieve if they are willing to open their property for short-term leasing for the entire calendar year and, ceteris paribus, the remainder of the market does not do so.

As hinted at above in the Likely versus Potential comparison and as indicated in the logistic regression models in Tables 2 and 4, host behavior may be an important determinant of achieving short-term preference.  Additionally, as the difference between Likely and Potential revenue suggests, many short-term hosts are not listing their properties full-time.  There may be, therefore, different types of hosts operating in the Airbnb market. From anecdotal evidence there are at least two types of hosts in the market: 1) Those that lease out their properties all the time in the name of profit maximization (Profit Seekers); and 2) Those that lease out their property very infrequently, for example when they are away on vacation (Opportunistic Sharers).

If these are the dominant two forms of hosts in the market, then a distribution of properties by the percent of time which they are blocked should be highly bi-modal with profit seekers on the left and opportunistic sharers on the right. The left hand panel of Figure 6 shows that this breakdown holds somewhat, with peaks of hosts at each end, but that there are also a considerable number of hosts who block their properties some (25\% to 75\%) of the time.

There is also a third potential host type, that of the host who utilizies multiple platforms to list their short-term property, of which Airbnb may just be one. A potential method to identify these hosts is to look at the number of individual periods of blocked days.  A host using multiple platforms will likely have many small periods of blocked days on Airbnb as they will block out those days in which reservations are made on other platforms. We test to see if this is a meaningful metric by plotting the total count of blocked periods versus the percentage of blocked days in the right panel of Figure 6. We do find many hosts with a large number of blocked periods, but, unfortunately, no true pattern emerges.

>>>>

---
title: "To Airbnb?: A look at Long vs Shorm Term Leasing"
subtitle: "Andy Krause and Gideon Aschwanden"
author: "Zillow and University of Melbourne"
date: "Revised: November 10th, 2018"
output:
  pdf_document: default
  html_document: default
header-includes: 
   - \usepackage{setspace}\doublespacing
   - \usepackage{caption}
   - \captionsetup[table]{skip=10pt}
geometry: margin=3.3cm
bibliography: abb.bib
editor_options: 
  chunk_output_type: console
---

@@

  library(tidyverse)
  library(sf)
  library(geosphere)
  library(ggmap)
  library(multidplyr)
  library(tmap)
  library(xtable)
  library(ROCR)
  library(xtable)
```

@@

## Abstract

@@

To address these questions and to provide some guidance to potential and current investors and owners, we examine the short- and long-term residential rental markets in Melbourne, Australia. Melbourne has a thriving short-term rental market with over 20,000 total properties listed on the Airbnb portal, the source of our short-term market data. To avoid confounding market dynamics caused by traditional vacation rentals (beach, ski and other recreation) we limit our study area to the urbanized areas of Melbourne. Additionally, for any individual property the choice between short and long term leasing may be influenced by individual financial situations (tax and mortgage financing) as well as non-financial concerns such as time availability and personal attachement to the property.  While we analyze individual properties, we acknowledge that there are many unobserved condiseration and, therefore, we generalize our findings to market segments, locations and host policy decisions.

@@

We begin by filtering observations by time. Our hypothetical leasing strategy decision is that of an investor who purchases a property on September 1, 2015 and must decide on traditional long-term rental versus an Airbnb-type, short-term rental approach to generating income. To determine the likely LTR nightly rate for any given property we use the observed rental transactions from September 1, 2014 to August 31, 2015. Taking the Airbnb or STR approach allows the owner to change rates over time and is subjected to daily changes in the market (supply and demand factors). In this simulated example we will use Airbnb data from September 1, 2015 to August 31, 2016 to represent the actual Airbnb market conditions over the time period in question.

@@

Both anecdotal evidence and Airbnb's own marketing campaigns suggest that there are at least two types of hosts in the market: 1) Those that lease out their properties all the time in the name of profit maximization (Profit Seekers); and 2) Those that lease out their property very infrequently, for example when they are away on vacation (Opportunistic Sharers).  For the purposes of this research we are focused on Profit Seekers; those owner who would be faced with deciding between a long- and short-term leasing strategy.

If these are the two dominant forms of hosts in the market, then a distribution of properties by the percent of time which they are blocked (not listed on the portal) should be highly bimodal with Profit Seekers in one group and Opportunistic Sharers in another. The left hand panel of Figure 1 shows that this breakdown holds somewhat, with peaks of hosts at each end, but that there are also a considerable number of hosts who block their properties some of the time (25\% to 75\%) .

Based on the data it appears there may also be a third potential host type, that of the host who utilizies multiple platforms to list their short-term property, of which Airbnb may just be one. A potential method to identify these hosts is to look at the number of individual periods of blocked days.  A host using multiple platforms will likely have many small periods of blocked days on Airbnb as they will block out those days in which reservations are made on other platforms. We test to see if this is a meaningful metric by plotting the total count of blocked periods versus the percentage of blocked days in the right panel of Figure 1. We do find many hosts with a large number of blocked periods, but, unfortunately, no true pattern emerges.

Attempts to divide the hosts up with unsupervised^[As there are no official host types we could not run more reliable supervised models.] clustering algorithms -- including the use of more variables -- failed to produce robust categories of hosts. As a more subjective, but implementable, approach we have divided the hosts up into three categories based on the original two variables considered: block percentage and number of blocked periods. Profit Seekers are those host who have blocked percentages less than 25\% of the time.  Opportunistic Sharers have blocked their properties at least 75\% of the time.  Within those two extremes, we consider users we consider as Unknown as they may be multiple plaform users (who are profit seeking), very opportunistic sharers or driven by some other, unknown motivation.  Profit Seekers make up 43\% of the hosts, with Opportunistic Sharers capturing 19\% and the remainder 'Unknown' category contains just over 38\% of all properties.  To best approximate the scenario considered in our research question, we keep only the Profit Seekers in our dataset. See Figure 2 for a breakdown of these group by the two variables.

<!-- ![Host Categories](C:/Code/research/airbnbMelbourne/images/host_scat.png) -->

@@

#### Rates and Rental Values

@@

#### Location and Submarkets

@@

We begin by eliminating any observations in the outer suburbs.  Very few short-term rentals were located in these areas and those that were have a much higher likelihood of being vacation homes; a group of properties uses that doesn't match well to our research question. While the Inner and Middle suburb designation is a useful starting place, we can imagine that a few different spatial features can influence the short-term market.  While there are many possible additional features, we consider two to be the most likely to influence rates or occupancy: 1) Proximity to beaches; and 2) Proximity to key tourist activities and events.

Using the two remaining broad submarkets (Inner and Middle) we have created a four sub-market system as a starting point for spatially analyzing short and long-term rentals in Melbourne:

1. City-Core (Select inner suburbs with tourist activities and near CBD)
2. City (Inner Suburbs, not Beach, not Core)
3. Suburban (Middle suburbs, not Beach)
4. Beach (Properties within 500m of Port Phillip Bay east of Yarra River)

@@

We assign these designations by: 1) Adding suburb designations to the properties; 2) Assigning submarkets 1-4 based on suburb location; and 3) Indicating proximity to beach and labeling as 'Beach' submarket.  This process is repeated for both the short- and long-term data. Finally, we filter our data to observations in suburbs that contain at least three STR and LTR observations.  The boundary of the suburbs meeting these criteria compose our study area (Figure 3). The location of the short- and long-term rentals, colored by submarket, are shown in Figures 4 and 5.

@@

To better visualize the relative frequency of the submarkets in the two tenure types we have broken down the data by submarket by property type by tenure type. A number of interesting observations are illustrated here (Figure 4). First, the vast majority of Airbnb units are apartments in city-core or city locations. As expected and as shown by the map (Figure 2), the long-term properties are more evenly spread across the metro region than the short-term properties.  Long-term houses are more peripheral than apartments, a result of urban development patterns and land economics. Beach properties are, relatively speaking, over-represented in the short-term market, especially in apartments, while suburban apartments, somewhat common in long-term markets, are nearly absent in the short-term market. Overall, the suburban Airbnb market is somewhat thin, later results focusing on these markets will need to be examined in light of this small sample size.

@@

@@

First, we must deal with those properties that were not listed on Airbnb for the entire 12-month period of our study.  Our data cleaning eliminated all properties that were listed for less than six months, however, it would be incorrect to compare the revenue of a property listed on Airbnb for seven months with the hypothetical annual revenue from a long-term rental.

Second, many of the short-term properties have a significant number of blocked days (up to 255), days in which the property was not offered on Airbnb.  Here too, there is a problem with comparing observed revenues from a property with a 20\% block rate with a hypothetical year long revenue stream, especially if the property was leased out on other portals during those blocked days.  And, even if we knew that the blocked days were simply due to the host wanting to use the property themselves during that time, this deduction in revenue should not necessarily apply to the hypothetical investor in our examined scenario.  In short, in either case, the actual observed short-term revenue would be biased on the low end compared to what a revenue maximizing owner would be likely able to accomplish.

As a result, we have calculated two separate revenues values for the short-term properties:

1. Actual ($REV_a$)}: Observed bookings multiplied by median nightly rate

2. Potential ($REV_p$)}: Observed bookings extrapolated to a full year plus imputed bookings during observed blocked days.  Our attempt here is to represent the potential revenue if the property had been listed all year long.  The imputed bookings are imputed by comparing the property specific performance during the non-blocked days to overall market performance and then applying that relative ratio to the overall market performance on the extrapolated and blocked days.

As the 'Actual' calculation systematically under counts the revenue from properties that were not on the Airbnb portal for the entire year we disregard it going forward and consider the 'Potential' revenue metric to be a better estimate of market performance. We chart the distribution of the Actual ($REV_a$) and Potential ($REV_p$) short-term revenues against a distribution of the observed long-term revenues for the long-term properties (Figure 7). Revenues from the long-term rentals are far more consistent, with the vast majority totaling between \$12,000 to \$30,000 per year.  Actual short-term revenues range from neglible to highs of greater than \$60,000, with many of the properties showing very limited (less than \$10,000) revenue. The range of the distribution for the Potential short-term revenue is similar, however, the distribution itself is shifted slightly to the right due to our time extrapolation and blocked days imputation.

While this simple analysis does not control for location and product differences, it does highlight the fact that some Airbnb owners are likely making greater revenues than in long-term rentals, but that most are likely not. The high number of very low yield Airbnb properties suggests that some owners may not be focused on maximizing revenue, but instead prefer the flexibility of short-term leasing.^[Do note that we have not considered any overall market demand issues when analyzing potential revenue.]

@@

In this section we compare net revenues between long- and short-term rental options.  As mentioned above, there are no direct matches between the datasets. In other words, we have not directly observed a property which has been in a long-term rental followed by short-term rental or vice versa.  We will, instead, make the comparisons between the two by imputing long-term rental revenues for the observed short-term properties and then comparing the measured net revenues of the short-term properties to their estimated imputed long-term revenues.

We make our imputations via an average from a two model ensemble. First we impute long term rental rates for the short methods with a hedonic price model. We use the following model specification:

@@

To limit the potential impacts of systematic bias from omitted variables or misspecification in our hedonic model on later analyses, we've also used a matched-pairs or clustering approach as the other half of our ensemble.  In this method, we first group properties based on structure type, bedroom and bathroom combination and then find the geographically nearest 7 long-term observations. For example, all 2-bed, 2-bath apartments are grouped.  For each short-term rental in this group we find the nearest 7 comparables (geographically), many of which are in the same building.  We compute the median of these seven comparables as our matched pairs estimation.

We then average the hedonic model and matched pairs prediction for each short term property to create an estimate of its long-term rental rate.

@@

Overall, 57.3%^[If we just look at actual revenue, ignoring blocked days and properties with less than a full 12 months of listing, the figure is 42.5%] of the 2,415 short-term rental properties in our sample were more profitable as a short-term rental than it would have been if rented long-term over the course of the year. While illustrative, a simple univariate analysis does not fully control for property characteristics or owner actions -- which are a key component of short-term rental success.  To better understand which factors are related to a higher likelihood of short-term preference, we develop a set of logistic regression models below.

@@

The results of all three models are shown in Table 2. For $PType$ Apartments are the reference category,  for $BedBath$, 1 Bed / 1Bath is the reference category, for $Submrkt$ the City-Core is the reference category and for $CancelPolicy$ Strict is the reference category.  As the additional variables are added, the model fit diagnostics steadily increase, suggesting that the increased complexity of the models adds to the quality of the fit.

In Model 1, we see that when controlling for structural characteristics, Houses have a significantly lower probability of being more profitable in short-term leasing than apartments. Additionally, 1 bed / 1 bath units show greater likelihood of short-term preference than 2 bed / 2 bath combinations. This could be due to the fact that 2 bed / 2 bath units may be very attractive in the long-term market as they are a useful product for two, non-family roomates to share. Larger units, 3 bedrooms plus, generally show no statistical difference in preference porbability to the 1 bed / 1 bath reference unit.

Adding the submarket variables (Model 2) shows that all areas have a much lower probability of short-term prefernce when compared to the City Core (the reference category) when controlling for structural characteristics. The largest difference is, unsurprisingly, between Suburban and City-Core properties. The coefficients for the structural characteristics show the same signs and relatively small changes in magnitude from Model 1 to Model 2, speaking to the robustness of those estimates.

Finally, in Model 3 we add in hostpolicies. Allowing a greater number of guests per bedroom results in a higher probability of short-term preference.  This variable may be operating as proxy for additional sleeping spaces which allow for a higher nightly rate, ceteris paribus, to be charged.  The greater the minimum stay, the lower the probability of short-term preference.  Longer minimum stays may equate to lower occupancy rates as one or two night travelers are exempted from the property.  Owners with longer minimum stays may be looking to minimize costs such as cleaning and administration.  Flexible cancellation policies result in a lower probability of short-term preference, while a Strict policy has a positive relationship with short-term preference. Strict cancellation may be a luxury available only for the best and most highly desired properties on the market and, within our model, may be acting as a proxy for some unobserved measure of quality or demand while a flexible cancellation may be signalling the opposite.

@@

  m1 <- as.data.frame(summary(pot_glms$str)$coefficients)
  m2 <- as.data.frame(summary(pot_glms$loc)$coefficients)
  m3 <- as.data.frame(summary(pot_glms$host)$coefficients)

@@

 ##  Create the variable section of table
 full.table <- as.data.frame(cbind(rep('x', length(m1c)), m1c, m2c, m3c))
 names(full.table) <- c('Variable', 'Model 1', "Model 2", "Model 3")
 full.table$Variable <- c('Intercept', 'House (Type)', '2 Bed & 1 Bath',
                          '2 Bed & 2 Bath', '3 Bed & 1 Bath', '3 Bed & 2 Bath',
                          '3 Bed & 3 Bath','4 Bed & 2 Bath','4 Bed & 3 Bath',
                          'City (Submarket)',
                          'Suburban (Submarket)',
                          'Beach (Submarket)', 'Guests per Bedroom',
                          'Minimum Stay', 'Flexible Cancel')

@@

  # Extract diagnostics
  str.dx <- logDx(pot_glms$str, str_tdf, str_tdf$pot_pref)
  sm.dx <- logDx(pot_glms$loc, str_tdf, str_tdf$pot_pref)
  host.dx <- logDx(pot_glms$host, str_tdf, str_tdf$pot_pref)

@@

  # Print
  print(x.full,
       include.rownames=FALSE,
       hline.after=c(-1,0,15,20,20),
       caption.placement='top')

@@

To test if the above results are influence by the assumptions we have made throughout this analysis, we perform a sensitivity test. The most influential assumptions that we make are in regards to; 1) unobserved costs of short term leasing (furnishings, additional maintenance, turnover, etc.) and that the properties in the two markets in our datasets are equivalent in terms of unobserved variables, such as quality and condition.  To approximate how these assumption may be influencing our results, we re-estimating our models under two new conditions:

1) **Missed Costs**:  Assume that there are extra costs for short term leasing that were not accounted for in our earlier analysis. Lowers relative preference of short-term option.

* Net short term revenues for short term rentals are systematically decreased 15%

2) **Uneven Sampling**: Assume that short term rental units are of a lower quality and condition than those in the long term market. Raises relative preference of short term option.

* Net long term revenues for short term rental are systematically decreased 15%

Table 3 shows the results of our sensitivity testing, with the original model results shown for comparison.  For the five variables with the strongest relationship in the original model -- House, City, Suburban, Guests Per Bedroom and Flexible Cancel -- the signs of the coefficient remain identical across the sensitivity tests and the variation in magnitudes are small, relatively speaking.   A few of the originally less significant variables do show relatively larger deviations in coefficient magnitude, but in general, the models appear to be stable across the sentivity tests.  Do note that the intercept terms does change across models, which would be expected as this represents the base probability of short-term preference which we'd expect to move in concert with our changing assumptions
